Sampling requirements to detect ecosystem change

### 2019-11-20 R/temporal-sampling.Rmd
Added theoretical results

We began with a theoretical exploration of the determiners

We modeled a simple AR-1 process with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal variability ($\sigma$), autocorrelation ($\phi$), the size of the shift ($\delta$), and a specific time when the shift occured ($\tau$):

$$
Model here
$$

@@

>>>>

We began with a theoretical exploration of the determiners of sampling required to detect a changepoint. We modeled a simple AR-1 process with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal variability ($\sigma$), autocorrelation ($\phi$), the size of the shift ($\delta$), and a specific time when the shift occured ($\tau$):

$$
\mbox{Model here - need pair of equations}
$$

@@

```{r theoretical_results, fig.cap='Theoretical results here',message=F,warning=F,echo=F}
# 4(or 6?) panel plot
## Example time series
## Changepoint detection vs number of samples - maybe just stick with regular sampling
## Number of samples required vs shift size, shift time, variance, autocorrelation

# Need to build function to output the time series
# Use purrr to try with a bunch of different parameters

source('R/uncertainty_tolerance.R')

build_time_series <- function(mu=0,sigma=5,phi=0.5,shift_size=25,shift_time=50,max_time=100){

  ar1 <- vector(mode = 'numeric',max_time)
  ar1[1] <- 100
  for (t in 2:max_time){
    if(t == shift_time){ar1[t-1]=(ar1[1]-shift_size)}
    ar1[t] <- ar1[t-1] + phi*rnorm(1,mean=mu,sd = sigma)
  }
  return(ar1) 
}

require(tidyverse)

par(mfrow=c(2,2))
# Sigma plot
params <- tibble(sigma=rep(seq(0.1,20,0.1),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 5, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(sigma))) %>%
  plot(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1)

# phi plot
params <- tibble(phi=rep(seq(-1,1,0.01),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(phi) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(phi))) %>%
  plot(ylab='Probability correct',xlab='phi',pch=16,cex.lab=1.3,las=1)

# shift_size plot
params <- tibble(shift_size=rep(seq(1,95,0.5),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_size) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(shift_size))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1)

# Shift timing plot
params <- tibble(shift_time=rep(seq(5,90,1),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_time) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_time/unique(shift_time))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1)

#params <- tibble(phi=rep(seq(-1,1,0.01),1),shift_size=rep(30,201))
#output <- params %>%
#  pmap(build_time_series)
#plot(output[[1]],type='l')
#points(output[[15]],type='l')
```
